Hierarchical Dispatching and Overriding

### 2017-06-17 Aplas17/sections/RelatedWork.tex
added references

\section{Related Work}

\subsection{Mainstream Multiple Inheritance Models}
Multiple inheritance is an useful feature in object-oriented
programming, although it is difficult to model and can 
cause various problems (e.g. the diamond problem~\cite{Sak89dis,Singh1995}).  There are many existing languages/models that support multiple 
inheritance. The Mixin model~\cite{} allows naming components 
that can be applied to various classes as reusable functionality units. However, the linearization (total ordering) of mixin 
inheritance cannot provide a satisfactory resolution in some cases and restricts the flexibility of mixin composition.

Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
strong line between units of reuse and object factories.  Traits act
as units of reuse, containing functionality code; while classes,
assembled from traits, act as object factories.\bruno{Should talk
  about work on traits that support renaming.}\bruno{I remember Yanlin
  had a presentation with alot of related work on multiple
  inheritance. Many of those seem to be missing here} Java 8
interfaces~\cite{} are closely related to traits: concrete method
implementations are allowed (via the \textbf{\texttt{default}}
keyword) inside interfaces.  The introduction of default methods opens
the gate for various flavors of multiple inheritance in Java.

@@

\noindent {\bf C++ model.}
C++ supports a very flexible inheritance model and allows programmers to choose either static dispatch or dynamic dispatch for method lookup.
It allows unintentional conflicts and uses static dispatch to resolve them, as discussed in Section~\ref{sec:overview}. For example, given the following code
\begin{lstlisting}[language=Java]
class A { public: void m() {cout << "MA" << endl;} };
class B { public: void m() {cout << "MB" << endl;} };
class C : public A, public B { 
	void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MA
}
\end{lstlisting}
The running result is $MB \; MA$, meaning that it uses static dispatch (looks at the static types) for method \lstinline|m()|. 
On calling \texttt{func(c)}, in spite that the dynamic type of \texttt{c} is class \texttt{C}, the method call still dispatches to 
\lstinline|A.m|.
However, we can alter the code a little bit with \textbf{\texttt{virtual}} methods and the result will be totally different:
\begin{lstlisting}[language=c++]
class A { public: virtual void m() {cout << "MA" << endl;} };
class B { public: virtual void m() {cout << "MB" << endl;} };
class C : public A, public B { 
    public: virtual void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MC
}
\end{lstlisting}
Now the running result will be $MB \; MC$. With virtual methods, dynamic dispatch is used and 
method lookup algorithm will find the most specific method definition of $m$, namely \lstinline|C.m| at this time.
Although C++ support this flexibility, dynamic dispatching on
conflicting methods is problematic.\bruno{why? give example of
  something that goes wrong?}
Furthermore, C++ does not support partial overrides compared to our
model.

>>>>

\section{Related Work}
Here we describe related work in four parts. We discuss from mainstream popular multiple 
inheritance models to several specific models (C++ and \csharp) which are closest to our work. Then we 
discuss similar techniques used in \self. Finally we discuss the foundation and related work of our formalization.

\subsection{Mainstream Multiple Inheritance Models}
Multiple inheritance is an useful feature in object-oriented
programming, although it is difficult to model and can 
cause various problems (e.g. the diamond problem~\cite{Sak89dis,Singh1995}).  
There are many existing languages/models that support multiple 
inheritance~\cite{ellis1990annotated,scala-overview,bracha90mixin,scharli03traits,malayeri2009cz,csharpdoc,Moon1986,Flatt1998,Ancona2003}. 
The Mixin model~\cite{bracha90mixin} allows naming components 
that can be applied to various classes as reusable functionality units. However, the linearization (total ordering) of mixin 
inheritance cannot provide a satisfactory resolution in some cases and restricts the flexibility of mixin composition. 
Scala traits~\cite{scala-overview} are in fact linearized mixins and hence have the same problem as mixins.

Simplifying the mixins approach, traits~\cite{scharli03traits,Ducasse:2006:TMF:1119479.1119483} draw a
strong line between units of reuse and object factories. Traits act
as units of reuse, containing functionality code; while classes,
assembled from traits, act as object factories. Java 8
interfaces are closely related to traits: concrete method
implementations are allowed (via the \textbf{\texttt{default}}
keyword) inside interfaces and there are also proposals such as FeatherTrait Java~\cite{Liquori08ftj} 
for extending Java with traits. There are extensions~\cite{reppy2006foundation,Reppy:2007:MT:2394758.2394784} to 
the original trait model with various advanced features such as \emph{renaming}. As discussed in Section~\ref{sec:overview},
the renaming feature helps to solve the unintentional method conflicts problem, however, it breaks structural subtyping.
\bruno{Should talk
  about work on traits that support renaming.}\bruno{I remember Yanlin
  had a presentation with alot of related work on multiple
  inheritance. Many of those seem to be missing here}

\begin{comment}
There are also proposals for extending Java with traits. For example, 
FeatherTrait Java (FTJ) [14] extends FJ [13] with statically-typed traits, 
adding trait-based inheritance in Java. Except for few, mostly syntac- tic details, 
their work can be emulated with Java 8 interfaces. There are also extensions 
to the original trait model, with operations (e.g. renaming [18], which breaks 
structural sub- typing) that default methods and interfaces cannot model.
\end{comment}

@@

\noindent {\bf C++ model.}
C++ supports a very flexible inheritance model and allows programmers to choose either static dispatch or dynamic dispatch for method lookup.
It allows unintentional conflicts and uses static dispatch to resolve them, as discussed in Section~\ref{sec:overview}. For example, given the following code
\begin{lstlisting}[language=Java]
class A { public: void m() {cout << "MA" << endl;} };
class B { public: void m() {cout << "MB" << endl;} };
class C : public A, public B { 
	void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MA
}
\end{lstlisting}
The running result is $MB \; MA$, meaning that it uses static dispatch (looks at the static types) for method \lstinline|m()|. 
On calling \texttt{func(c)}, in spite that the dynamic type of \texttt{c} is class \texttt{C}, the method call still dispatches to 
\lstinline|A.m|.
However, we can alter the code a little bit with \textbf{\texttt{virtual}} methods and the result will be totally different:
\begin{lstlisting}[language=c++]
class A { public: virtual void m() {cout << "MA" << endl;} };
class B { public: virtual void m() {cout << "MB" << endl;} };
class C : public A, public B { 
    public: virtual void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MC
}
\end{lstlisting}
Now the running result will be $MB \; MC$. With virtual methods, dynamic dispatch is used and 
method lookup algorithm will find the most specific method definition of $m$, namely \lstinline|C.m| at this time.
Although C++ support this flexibility, dynamic dispatching on
unintentional conflicting methods is problematic, as discussed in Section~\ref{sec:overview}.\bruno{why? give example of
  something that goes wrong?}
Furthermore, C++ does not support partial overrides compared to our
model.
